Manuscript Draft\\ \vspace{2cm} What is long-distance dispersal? And a taxonomy of dispersal events

### 2016-06-28 Manuscript/MS- What is LDD.tex
First draft finished.

\textbf{Key words}: ***\\

@@

The frugivore assemblage of \textit{P. mahaleb} is composed by a diversified set of animal species spanning a wide size range, ca. 12-14000 g in body mass. We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}. If the results for \textit{P. mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events, and large-bodied species with a disproportionate contribution to LDD events. Both components of this sort of diplochorous \citep{vanderWall:2004hv} dispersal system are very frequent in fleshy-fruited plants with diversified frugivore assemblages \citep{Galetti:2013jv}. In such cases, small-bodied frugivores largely contribute the short-distance dispersal key to support \textit{in situ} recruitment and population persistence. Yet the large-bodied frugivores distinctly contribute LDD events that sustain the connectivity of metapopulation scenarios \citep{Urban:2001}. Yet, as shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools.

A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal: i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisititation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}. Yet these large-bodied frugivores are crucial for both $LDD_{loc}$ and $LDD_{ss}$, given that extensive movement patterns and extremely large foraging ranges may frequently contribute dissemination beyond distance thresholds defined with either spatial landscape or genetic references. Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep{Westcott:2005,Bueno:2013cg}.

As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a specific case of extreme events \citep{Garcia:2017aa} consistently associated with large-sized frugivores, yet including also medium-sized and highly efficient frugivorous bird species.

- Large-bodied frugivores and LDD.  Discuss lb frugivores not always contributing LDD events: Pratt, tapirs
- Loss of large-bodied frugivores, consequences.

@@

Pollen and seed dispersal in plants are essentially spatially-structured processes for which the outcomes of interactions with dispersal vectors is intimately linked to landscape features. Given this mechanistic link between the features of the vector and the environments where its displacement occurs \citep{Nathan:2008fx}, consideration of landscape is key to understand the consequences of LDD events. Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow .

>>>>

\textbf{Key words}: dispersal, frugivory, plant-animal interactions, pollination, seed dispersal\\

@@

The frugivore assemblage of \textit{P. mahaleb} is composed by a diversified set of animal species spanning a wide size range, ca. 12-14000 g in body mass. We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}. If the results for \textit{P. mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events, and large-bodied species with a disproportionate contribution to LDD events. Both components of this sort of diplochorous \citep{vanderWall:2004hv} dispersal system are very frequent in fleshy-fruited plants with diversified frugivore assemblages \citep{Galetti:2013jv}. In such cases, small-bodied frugivores largely contribute the short-distance dispersal key to support \textit{in situ} recruitment and population persistence. Yet the large-bodied frugivores distinctly contribute LDD events that sustain the connectivity of metapopulation scenarios \citep{Urban:2001}. As shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools. For example, local, within-population, dispersal events may vary enormously in terms of genetic effects and local structuring of the genetic pools depending on whether they specifically contribute $SDD_{loc}$ or instead, $LDD_{loc}$. Note that only the latter actually contribute erasing any form of local genetic structure by contributing to increased genetic neigborhoods.

A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal because of, i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisitation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}. Yet these large-bodied frugivores are crucial for both $LDD_{loc}$ and $LDD_{ss}$, given that extensive movement patterns and extremely large foraging ranges may frequently contribute dissemination beyond distance thresholds defined with either spatial landscape or genetic references. Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep[e.g., ][]{Westcott:2005,Bueno:2013cg,Morales:2013dg,Carlo:2013vd}. In addition, medium-sized birds such as thrushes (\textit{Turdus} spp.) can contribute substantial $LDD_{loc}$ events, i.e., local LDD events contributing to erase local population genetic structuring, effectively increasing the size of genetic neighborhoods. In the case of \textit{P. mahaleb} up to 55.49\% of their dispersal events are $LDD_{loc}$ events. These birds are efficient seed dispersers of \textit{P. mahaleb} and other fleshy-fruited species \citep{Snow:1988ve,Jordano:2000ft,Carlo:2013vd}, also showing significant contributions of $LDD_{ss}$ events.

As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a specific case of extreme events \citep{Garcia:2017aa} consistently associated with large-sized frugivores, yet including also medium-sized and highly efficient frugivorous bird species. Robustly characterizing the expected frequencies and extent of those extreme events would be crucial to properly assess the functional role of frugivores and the full range of influences (demographic, genetic) in plant populations.

@@

Pollen and seed dispersal in plants are essentially spatially-structured processes for which the outcomes of interactions with dispersal vectors is intimately linked to landscape features. Given this mechanistic link between the features of the vector and the environments where its displacement occurs \citep{Nathan:2008fx}, consideration of landscape is key to understand the consequences of LDD events. Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow in complex landscapes \citep{Sork:1999}. Recent evidences point out that the selective extinction of large-bodied frugivores may significantly impact plant populations dependent on frugivores both in terms  of recruitment \citep{Traveset:2012he,PerezMendez:2015hya} and genetic connectivity \citep{Perez-Mendez:2016dz}. Frugivore downsizing represents a lasting challenge for the collapse of seed dispersal processes where $LDD_{ss}$ events are crucial for population persistence and the cohesion of fragmented populations within metapopulation scenarios.

We advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research. A crucial aspect would be to effectively associate the role of individual frugivore species to specific dispersal outcomes, by identifying the actual disperser contributing a dissemination event \citep{GonzalezVaro:2014ij} and simultaneously characterizing the source maternal plant \citep{Jordano:2002tr}.

The actual challenges to properly characterize the typologies of LDD events outlined in Table 1 will probably persist. We need more efficient quantitative approaches to assess these infrequent events, that occur over enormous spatial scales and that need to be documented with sample sizes sufficient to facilitate modeling efforts and robust statistical inferences. These are not trivial difficulties given the urgency to assess how forest loss, defaunation, genetic purging due to logging, etc., alter plant populations.
